### 2023-03-24 paper/proof.tex
initial commit

\section{Assumptions}\label{sec:assumptions}
This section is a draft, possibly contains mistakes
Assumptions on the model required for deriving regret bounds, in the spirit of \cite{curiCombining2021}. \cref{ass:lipschitz-f} and \cref{ass:lipschitz-gp} are copied from \cite{curiCombining2021}, whereas \cref{ass:calibrated} has been modified to accommodate the mean and variance changing during the episode.
\begin{assumption}\label{ass:lipschitz-f}
The true underlying function $f$ is $L_f$-Lipschitz and the elements of the noise vector $w_n$ are $\sigma$-sub-Gaussian.
\end{assumption}
\begin{assumption}[Well-calibrated]\label{ass:calibrated}
Assume that for any $\delta >0$ $\exists$ a set of sequences $\{\beta_{t, i}\}_{t>0, i=0}^{episode_length}$ such that the model mean and variance functions $\mu_{t, i}$, $\sigma_{t, i}$ satisfy for all $(s, a) \in \mathcal{S} \times \mathcal{A}$ the upper bound
\begin{equation}\label{eq:calibrated}
\text{Pr}(\| f(s, a) - \mu_{t, i}(s, a) \| \leq \beta_{t, i}\sigma_{t, i}) \geq 1 - \delta
\end{equation}
\end{assumption}

\begin{assumption}\label{ass:lipschitz-gp}
The mean and variance functions are lipschitz
\end{assumption}
\section{Proof of regret bounds}
This section is a draft, possibly contains mistakes
Plan for the proof:
\begin{itemize}
\item Following Appendix D in \cite{curiCombining2021}, derive a regret bound for the case where the model $\mu_t$ and $\sigma_t$ can be updated within the episode, if assumptions in \cref{sec:assumptions} are satisfied.
\item Show that sparse GPs satisfy \cref{ass:calibrated}
\end{itemize}