### 2023-05-16 paper/main.tex
rewrite 3 points

% !TeX spellcheck = en_GB
\documentclass{article}

% Pass options to natbib
\PassOptionsToPackage{numbers, compress}{natbib}

% NeurIPS packages
\usepackage[]{neurips_2023}
\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
%\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{xcolor}         % colors

% Redefine paragraph to be tighter
\renewcommand{\paragraph}[1]{{\bf #1}~~}

% Array/table packages
\usepackage{tabularx}
\usepackage{array,multirow}
\usepackage{colortbl}
\newcommand{\PreserveBackslash}[1]{\let\temp=\\#1\let\\=\temp}
\newcolumntype{C}[1]{>{\PreserveBackslash\centering}p{#1}}
\newlength{\tblw}

% Latin
\usepackage{xspace}
\newcommand{\eg}{\textit{e.g.\@}\xspace}
\newcommand{\ie}{\textit{i.e.\@}\xspace}
\newcommand{\cf}{\textit{cf.\@}\xspace}
\newcommand{\etc}{\textit{etc.\@}\xspace}
\newcommand{\etal}{\textit{et~al.\@}\xspace}

% Our method
\newcommand{\our}{\textsc{sfr}\xspace}

% Tikz
\usepackage{tikz}
\usepackage{pgfplots}
\usetikzlibrary{patterns}
\usetikzlibrary{decorations,backgrounds,arrows.meta,calc}
\usetikzlibrary{shapes,arrows,positioning}

% Appendix/supplement title
\newcommand{\nipstitle}[1]{{%
    % rules for title box at top and bottom
    \def\toptitlebar{\hrule height4pt \vskip .25in \vskip -\parskip} 
    \def\bottomtitlebar{\vskip .29in \vskip -\parskip \hrule height1pt \vskip .09in} 
    \phantomsection\hsize\textwidth\linewidth\hsize%
    \vskip 0.1in%
    \toptitlebar%
    \begin{minipage}{\textwidth}%
        \centering{\LARGE\bf #1\par}%
    \end{minipage}%
    \bottomtitlebar%
    \addcontentsline{toc}{section}{#1}%
}}

% Bibliography
%\usepackage[maxcitenames=1, maxbibnames=4, doi=false, isbn=false, eprint=true, backend=bibtex, hyperref=true, url=false, style=authoryear-comp]{biblatex}
%\addbibresource{zotero-library.bib}
% \addbibresource{paper/zotero-library.bib}

% Let's use good old bibtex instead

% Figure customization: Tight legend box
\pgfplotsset{every axis/.append style={
		legend style={inner xsep=1pt, inner ysep=0.5pt, nodes={inner sep=1pt, text depth=0.1em},draw=none,fill=none}
}}

% Our packages
\usepackage{todonotes}
\usepackage[colorlinks=true,linkcolor=blue,allcolors=blue]{hyperref}
\usepackage{amsmath}
\usepackage{bm}
\usepackage{algpseudocode}
\usepackage{algorithm}
\usepackage{derivative}

\usepackage{tikz,pgfplots}
\usepackage{subcaption}
\usetikzlibrary{}

\input{aidans-utils.tex}

% Short section names etc
% This must be imported last!
%\usepackage{cleveref}
\usepackage[capitalise,nameinlink]{cleveref}
\crefname{section}{Sec.}{Secs.}
\crefname{algorithm}{Alg.}{Algs.}
\crefname{appendix}{App.}{Apps.}
\crefname{definition}{Def.}{Defs.}
\crefname{table}{Table}{Tables}

% Config for Arno's awesome TikZ plotting stuff
\newlength{\figurewidth}
\newlength{\figureheight}

% Variables
\newcommand{\state}{\ensuremath{\mathbf{s}}}
\newcommand{\action}{\ensuremath{\mathbf{a}}}
\newcommand{\noise}{\ensuremath{\bm\epsilon}}
\newcommand{\discount}{\ensuremath{\gamma}}
\newcommand{\inducingInput}{\ensuremath{\mathbf{Z}}}
\newcommand{\inducingVariable}{\ensuremath{\mathbf{u}}}
\newcommand{\dataset}{\ensuremath{\mathcal{D}}}
\newcommand{\dualParam}[1]{\ensuremath{\bm{\lambda}_{#1}}}
\newcommand{\meanParam}[1]{\ensuremath{\bm{\mu}_{#1}}}

% Indexes
\newcommand{\horizon}{\ensuremath{h}}
\newcommand{\Horizon}{\ensuremath{H}}
\newcommand{\numDataNew}{\ensuremath{N^{\text{new}}}}
\newcommand{\numDataOld}{\ensuremath{N^{\text{old}}}}
\newcommand{\numInducing}{\ensuremath{M}}

% Domains
\newcommand{\stateDomain}{\ensuremath{\mathcal{S}}}
\newcommand{\actionDomain}{\ensuremath{\mathcal{A}}}
\newcommand{\inputDomain}{\ensuremath{\mathbb{R}^{D}}}
\newcommand{\outputDomain}{\ensuremath{\mathbb{R}^{C}}}
\newcommand{\policyDomain}{\ensuremath{\Pi}}

% Functions
\newcommand{\rewardFn}{\ensuremath{r}}
\newcommand{\transitionFn}{\ensuremath{f}}
\newcommand{\latentFn}{\ensuremath{f}}

\newcommand{\optimisticTransition}{\ensuremath{\hat{f}}}
\newcommand{\optimisticTransitionMean}{\ensuremath{\mu_{\optimisticTransition}}}
\newcommand{\optimisticTransitionCov}{\ensuremath{\mu_{\optimisticTransition}}}
\newcommand{\optimisticTransitionSet}{\ensuremath{\mathcal{M}}}

% Parameters
% \newcommand{\weights}{\ensuremath{\bm\phi}}
\newcommand{\weights}{\ensuremath{\mathbf{w}}}
\newcommand{\valueFnParams}{\ensuremath{\psi}}
\newcommand{\policyParams}{\ensuremath{\theta}}

% Networks
\newcommand{\transitionFnWithParams}{\ensuremath{\transitionFn_{\weights}}}
\newcommand{\valueFn}{\ensuremath{\mathbf{Q}}}
\newcommand{\stateValueFn}{\ensuremath{\mathbf{V}}}
% \newcommand{\valueFn}{\ensuremath{\mathbf{Q}_{\valueFnParams}}}
\newcommand{\policy}{\ensuremath{\pi}}
\newcommand{\pPolicy}{\ensuremath{\pi_{\policyParams}}}

% Packages for bold math
\usepackage{bm}
\newcommand{\mathbold}[1]{\bm{#1}}
\newcommand{\mbf}[1]{\mathbf{#1}}
\renewcommand{\mid}{\,|\,}

% Math Macros
\newcommand{\MB}{\mbf{B}}
\newcommand{\MC}{\mbf{C}}
\newcommand{\MZ}{\mbf{Z}}
\newcommand{\MV}{\mbf{V}}
\newcommand{\MX}{\mbf{X}}
\newcommand{\MA}{\mbf{A}}
\newcommand{\MK}{\mbf{K}}
\newcommand{\MI}{\mbf{I}}
\newcommand{\MH}{\mbf{H}}
\newcommand{\T}{\top}
\newcommand{\vzeros}{\mbf{0}}
\newcommand{\vtheta}[0]{\mathbold{\theta}}
\newcommand{\valpha}[0]{\mathbold{\alpha}}
\newcommand{\vkappa}[0]{\mathbold{\kappa}}
\newcommand{\vbeta}[0]{\mathbold{\beta}}
\newcommand{\MBeta}[0]{\mathbold{B}}
\newcommand{\vlambda}[0]{\mathbold{\lambda}}
\newcommand{\diag}{\text{{diag}}}

\newcommand{\vm}{\mbf{m}}
\newcommand{\vz}{\mbf{z}}
\newcommand{\vf}{\mbf{f}}
\newcommand{\vu}{\mbf{u}}
\newcommand{\vx}{\mbf{x}}
\newcommand{\vy}{\mbf{y}}
\newcommand{\vw}{\mbf{w}}
\newcommand{\va}{\mbf{a}}

\newcommand{\Jac}[2]{\mathcal{J}_{#1}(#2)}
\newcommand{\JacT}[2]{\mathcal{J}_{#1}^\top(#2)}

\newcommand{\GP}{\mathcal{GP}}
\newcommand{\KL}[2]{\mathrm{D}_\textrm{KL} \dbar*{#1}{#2}}
\newcommand{\MKzz}{\mbf{K}_{\mbf{z}\mbf{z}}}
\newcommand{\MKxx}{\mbf{K}_{\mbf{x}\mbf{x}}}
\newcommand{\MKzx}{\mbf{K}_{\mbf{z}\mbf{x}}}
\newcommand{\MKxz}{\mbf{K}_{\mbf{x}\mbf{z}}}
\newcommand{\vkzi}{\mbf{k}_{\mbf{z}i}}
\newcommand{\vkzs}{\mbf{k}_{\mbf{z}i}}
\newcommand{\vk}{\mbf{k}}
\newcommand{\MLambda}[0]{\mathbold{\Lambda}}
\newcommand{\MSigma}[0]{\mathbold{\Sigma}}
\definecolor{matplotlib-blue}{HTML}{1f77b4}
\newcommand{\N}{\mathrm{N}}
%\newcommand{\R}{\mathrm{R}}
\newcommand{\myexpect}{\mathbb{E}}

\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}
\newcommand{\Norm}{\mathcal{N}}

%\title{Investigatin Uncertainty Quantification in Model-based Reinforcement Learning}
% \title{Model-based Reinforcement Learning with Fast Posterior Updates}
%\title{Sequential Decision-Making under Uncertainty with Big Data}
% \title{Neural Network to Vatiational Sparse Gaussian Process: For Adaptive Exploration}
% \title{Neural Network to Sparse Variational Gaussian Process: For Updates in Sequential Decision Making}
% \title{Adapting Neural Networks to New Data For Updates in Sequential Decision Making via Gaussian Processes}
% \title{Converting Neural Networks to Gaussian Processes for Sequential Decision-Making Under Uncertainty}
%\title{Sparse Function Space Representation of Neural Networks for Exploration and Retention}
%\title{Sparse Function-space Neural Networks}
\title{Sparse Function-space Representation \\ of Neural Networks}% for Adaptation and Retention}
\author{%
  Aidan Scannell\textsuperscript{\star} \\
  Aalto University \\
  Finnish Center for Artificial Intelligence \\
  \texttt{aidan.scannell@aalto.fi}
  \And
  Riccardo Mereu\textsuperscript{\star} \\
  Aalto University\\
  \texttt{riccardo.mereu@aalto.fi}
  \And
  Paul Chang \\
  Aalto University\\
  \texttt{paul.chang@aalto.fi}
  \And
  Ella Tamir \\
  Aalto University\\
  \texttt{ella.tamir@aalto.fi}
  \And
  Joni Pajarinen \\
  Aalto University\\
  \texttt{joni.pajarinen@aalto.fi}
  \And
  Arno Solin \\
  Aalto University\\
  \texttt{arno.solin@aalto.fi}
}

\begin{document}

\maketitle

\begin{abstract}
% OLDER VERSION
%Sequential learning paradigms such as Continual Learning (CL) or Reinforcement Learning (RL) pose a challenge for gradient-based deep learning techniques as they struggle to incorporate new data and retain previous knowledge. Existing methods for converting neural networks from weight to function space allow a probabilistic treatment of the distribution over the function learned by the neural networks but are computationally expensive. We propose a method that converts a neural network to a low-rank functional representation as a sparse Gaussian process. With this approach, we can build a compact representation of the function encoded by the neural network that can replace previous data in continual settings and be used for fast adaptation in RL, avoiding full retraining of the model. 
%
% Rewrite on 2023-05-10
%Deep neural networks are known to lack uncertainty estimates, struggle to incorporate new data, and fail to retain previous knowledge. We present a method that mitigates these issues by transforming a weight-space neural network to a low-rank function-space representation, via the so-called dual parameters. In contrast to previous work, we model the joint distribution across the entire data set rather than a subset. This offers a compact and principled way of capturing uncertainty and enables us to incorporate new data without retraining whilst retaining predictive performance. We demonstrate the proposed approach for quantifying uncertainty in supervised learning and maintaining a compact representation in sequential learning.\looseness-1

Deep neural networks are known to lack uncertainty estimates, struggle to incorporate new data, and suffer from catastrophic forgetting. We present a method that mitigates these issues by converting neural networks from weight-space to a low-rank function-space representation, via the so-called dual parameters. In contrast to previous work, our sparse representation captures the joint distribution over the entire data set, rather than only over a subset. This offers a compact and principled way of capturing uncertainty and enables us to incorporate new data without retraining whilst retaining predictive performance. We demonstrate the proposed approach for quantifying uncertainty in supervised learning and maintaining an expressive functional representation for sequential learning.\looseness-1

\end{abstract}

%, maintaining a summary representation in continual learning,

\section{Introduction}
\label{sec:intro}
%
Deep learning \cite{goodfellow2016deep} has become the cornerstone of contemporary artificial intelligence, proving remarkably effective in tackling supervised and unsupervised learning tasks in the {\em large data}, {\em offline}, and {\em gradient-based training} regime. Despite its success, gradient-based learning techniques exhibit limitations. Firstly, how can we efficiently quantify uncertainty without resorting to expensive and hard-to-interpret sampling in the model's weight-space? Secondly, how to update the weights of an already trained model with new batches of data without compromising the performance on past data? These questions become central when applied to sequential learning paradigms, such as continual learning (CL, \citep{parisi2019continual, de2021continual}) and reinforcement learning (RL, \cite{sutton2018reinforcement}). In CL, access to the previous data is lost and the challenge is retaining a compact representation of the problem and avoiding forgetting over the life-long learning horizon~\cite{mccloskey1989catastrophic}. Similarly, in RL, the model must adapt to environmental observations through exploration, while leveraging prediction uncertainties to assess potential future paths.\looseness-1

%Secondly, how to retain information from previous tasks whilst learning new tasks where the tasks are such as Continual Learning (CL, \cite{de2021continual})
%Thirdly and differently from the CL problem is that given some data from the same distribution

%Current state of affairs 
Recent techniques (\eg, \cite{ritter2018kfac,khan2019approximate,daxberger2021laplace,fortuin2021bayesian,immer2021scalable}) show that converting trained neural networks to a Bayesian neural network through the Laplace-GGN can provide uncertainty without sacrificing additional training cost \cite{foong2019between}. Furthermore, the resultant weight space posterior can be converted to the function space as shown \cite{khan2019approximate, immer2021improving}. The function space allows for a principled mathematical approach for analyzing the behaviour \cite{cho2009kernel,meronen2020stationary}, performing probabilistic inference \cite{khan2019approximate}, and quantifying uncertainty in neural networks \cite{foong2019between}. These methods rely on the linearization of the neural network and the resultant NTK \cite{jacot2018neural}. The result function space neural network is characterized by its first two moment functions, a mean function and covariance function (or kernel)---defining a Gaussian process (GP, \cite{rasmussen2006gaussian}). GPs provide a widely-used probabilistic toolkit with principled uncertainty estimates, and they serve as a standard surrogate model for Bayesian optimization \citep{garnett_bayesoptbook_2022} and are effective in model-based reinforcement learning \citep{deisenroth2011pilco} with theoretical guarantees on regret bounds \citep{srinivas2009gaussian}. 
%Yet many problems lie in high dimensional input space; for example, images are where GPs cannot learn representations. In such scenarios such as in many reinforcement learning environments neural networks are used as the surrogate model. However, uncertainty is still essential to ensure effective exploration for sequential algorithms. Successful approaches have attempted to blend neural networks with uncertainty estimates around predictions, allowing for sophisticated exploration strategies. However, there has been limited use of hybrid models that possess the feature representation ability of neural networks but also attractive the properties of GPs, such a hybrid method we propose in this paper.

%Need for adaptive learning methods + failures with current methods
In this paper, we demonstrate that any neural network can be represented in terms of the so-called `dual' parameters of a GP~\cite{csato2002sparse, adam2021dual, chang2020fast}. In contrast to previous work that utilizes subsets \cite{immer2021scalable}, this parameterization allows for the consolidation of contributions from {\em all} training data into a sparse representation. Crucially, the resulting GP now predicts in the same space as the originally trained neural network, a feature not present in previous approaches, while still avoiding the notorious cubic complexity of vanilla GPs. Through the dual parameter formulation, we establish a connection between the neural network, the resulting full GP, and a sparse approximation similar to variational sparse GPs~\cite{titsias2009variational}. Using the sparse GP we can then take advantage of the property of dual parameters that allows for fast conditioning recently shown in \cite{chang2022fantasizing}. Thus we are able to avoid retraining and condition new data in to our model (see \cref{fig:teaser}). The resulting \our method can be applied in any sequential decision-making scenario where retraining is not feasible.

%Need uncertainty and adaptive methods
%Dual formulation in GPs space solves this
%Talk about planning and exploration in RL.

The contributions of this paper are:
%
{\em (i)}~We introduce \our, a new approach for sparse functional representation of the neural network.
{\em (ii)}~We demonstrate that, despite its sparsity, our method effectively captures predictive uncertainty, provides means of updating the model post-training, and gives a a compact representation suitable for continual learning.
{\em (iii)}~We provide extensive experiments for showcasing our approach and demonstrate significance and applicability across supervised, continual, and reinforcement learning, aiming to stimulate future use of the approach.

%List the contributions.
%The contributions of the paper our is as follows:
%\begin{itemize}
%\item We show how to take a trained neural network and convert it to a dual sparse GP. We are able to do this without retraining a variational objective for the Sparse GP. Our sparse GP uses the variational formulation, and thus gives better uncertainty estimates than other no variational sparse approaches used previously.
%\item The sparse GP now gives us a compact representation of our parameters in the function space. We can therefore take advantage of the dual parameters formulation for fast conditioning of new data in to our posterior avoiding retraining of the neural network. Crucially this allows for fast adaptation of models that our used in sequential decision making. We show how this is effective in the planning stage of model-based reinforcement exploration.
%\end{itemize}

%
%
%\begin{itemize}
%  \item Many real-world problems require learning-based systems that can adapt to new data.
%  \begin{itemize}
%    % \item For example, in domains such as robotics and healthcare,
%    \item For example, when controlling robots in non-stationary environments it is important for the robot to adapt to the changing dynamics.
%    \item However neural networks rely upon gradient-based optimisation.
%    \item Uncertainty can be used to improve sample efficiency via targeted exploration.
%    \item Uncertainty can be used to handle risk in decision making.
%  \end{itemize}
%
%  \item Gaussian processes can easily adapt to new data and they offer well-calibrated uncertainty estimates.
%  \begin{itemize}
%    \item However, they don't scale to high-dimensional and large data sets.
%  \end{itemize}
%  \item
%  \begin{itemize}
%    \item
%  \end{itemize}
%\end{itemize}

\begin{figure}[t!]
  \centering\scriptsize
  % Figure options
  \pgfplotsset{axis on top,scale only axis,width=\figurewidth,height=\figureheight, ylabel near ticks,ylabel style={yshift=-2pt},y tick label style={rotate=90},legend style={nodes={scale=1., transform shape}},tick label style={font=\tiny,scale=1}}
  \pgfplotsset{xlabel={Input, $x$},axis line style={rounded corners=2pt}}
  % Set figure 
  \setlength{\figurewidth}{.28\textwidth}
  \setlength{\figureheight}{\figurewidth}
  %
  \def\inducing{\large Sparse inducing points}
  %
  \begin{subfigure}[c]{.34\textwidth}
    \raggedleft
    \pgfplotsset{ylabel={Output, $y$}}
    \input{./fig/regression-nn.tex}%
  \end{subfigure}
  \hfill  
  \begin{subfigure}[c]{.01\textwidth}
    \centering
    \tikz[overlay,remember picture]\node(p0){};
  \end{subfigure}  
  \hfill
  \begin{subfigure}[c]{.28\textwidth}
    \raggedleft
    \pgfplotsset{yticklabels={,,},ytick={\empty}}
    \input{./fig/regression-nn2svgp.tex}%
  \end{subfigure}
  \hfill  
  \begin{subfigure}[c]{.01\textwidth}
    \centering
    \tikz[overlay,remember picture]\node(p1){};
  \end{subfigure}  
  \hfill
  \begin{subfigure}[c]{.28\textwidth}
    \raggedleft
    \pgfplotsset{yticklabels={,,},ytick={\empty}}        
    \input{./fig/regression-update.tex}%
  \end{subfigure}
  \caption{\textbf{Regression example on an MLP with two hidden layers.} Left:~Predictions from the trained neural network. Middle:~Our approach summarizes all the training data with the help of an iducing set of points. The model captures the predictive mean and uncertainty, and (right) makes it possible to update the model with new data without degrading previous performance.}
  \label{fig:teaser} 
  % 
  \begin{tikzpicture}[remember picture,overlay]
    % Arrow style
    \tikzstyle{myarrow} = [draw=black!80, single arrow, minimum height=14mm, minimum width=2pt, single arrow head extend=4pt, fill=black!80, anchor=center, rotate=0, inner sep=5pt, rounded corners=1pt]
    % Arrows
    \node[myarrow] (p0-arr) at ($(p0) + (1em,1.5em)$) {};
    \node[myarrow] (p1-arr) at ($(p1) + (1em,1.5em)$) {};
    % Arrow labels
    \node[font=\scriptsize\sc,color=white] at (p0-arr) {\our};
    \node[font=\scriptsize\sc,color=white] at (p1-arr) {new data};   
  \end{tikzpicture}
\end{figure}

\subsection{Related work}
\label{sec:related}
%
% General Bayesian deep learning
\textbf{Bayesian deep learning}
Probabilistic methods in deep learning~\cite{Wilson:ensembles,neal1995bayesian} have recently gained increasing attention in the machine learning community as a means for uncertainty quantification (\eg, \cite{kendall2017what,wilson2020bayes}) and model selection (\eg,~\cite{immer2021scalable,antoran2022marginal}) with advancements in prior specification (\eg, \cite{cho2009kernel,meronen2020stationary,meronen2021periodic,fortuin2021bayesian,nalisnick2018do}) and efficient approximate inference under the specified model.
Calculating the posterior distribution of a Bayesian neural network is usually intractable, and approximate inference techniques need to be used, such as variational inference \cite{blei2017variational}, deep ensembles \cite{lakshminarayanan2017simple}, MC dropout \cite{gal2016dropout}, or Laplace approximation \cite{ritter2018kfac,kristiadi2020being,immer2021improving}---each with its own set of strengths and weaknesses.

%\textbf{Function-space methods}
%One common approach for uncertainty with neural networks is a Laplace-GGN approximation \citep{daxberger2021laplace}, which takes a trained neural network and linearises it around the optimal weights. The Hessian can be efficiently approximated using the generalized Gauss--Newton approximation (GGN)~\cite{botev2017practical} but typically involves a cubic scaling in the number of parameters, in practice a further approximation such as the Kronecker factorisation \cite{martens2015optimizing,ritter2018kfac} is needed. The resulting linear model can be used to obtain uncertainty estimates and refine the neural network predictions \citep{immer2021scalable}, and is linear with respect to the weights. Therefore, as shown in \cite{immer2021scalable, khan2019approximate,maddox2021fast}, the linear model can be converted to a GP, which is then used to obtain the uncertainty estimates.

% Introduce the laplace GGN + shortcomings
\textbf{Function-space methods}
Function-space perspectives on uncertainty in neural networks often use a Laplace-GGN approximation \citep{daxberger2021laplace}, linearizing a trained neural network around optimal weights. The Hessian can be efficiently approximated using the generalized Gauss--Newton approximation (GGN, \cite{botev2017practical}). While efficient, it suffers from cubic scaling in parameter count, necessitating approximations like Kronecker factorisation \cite{martens2015optimizing, ritter2018kfac}. This linear model, linear in the weights, refines predictions and provides uncertainty estimates \citep{immer2021scalable}. This model has a convenient interpretation as a GP \cite{immer2021scalable, khan2019approximate, maddox2021fast}.
%
However, the GP's cubic $\mathcal{O}(n^3)$ scaling with data points $n$ requires costly approximations, often resorting to using data subsets \cite{immer2021scalable}. In the GP community, sparse approximations have mitigated this scaling issue (\eg, \cite{titsias2009variational,hensman2013gaussian}), but combining neural network linearization with sparse methods remains unclear without resorting to subset methods. Our work addresses this by utilizing the GP's dual parameters \cite{csato2002sparse}, previously applied to non-conjugate likelihood models \cite{adam2021dual}.

%wang2019exact
%
%However, the GP introduces a cubic $\mathcal{O}(N^3)$ scaling in the number of data points $N$. Previous approaches apply crude approximations to alleviate the computational costs, \eg, by considering only a subset of the training data \cite{immer2021scalable}. In the GP community, the scaling problem is not seen as an issue any more due to efficient and stochastic methods for sparse approximations (\eg, \cite{hensman2013gaussian,wang2019exact}). However, it is not entirely clear how to combine the linearization of the neural network with sparse methods \cite{titsias2009variational} without falling back to crude subset methods. We tackle this problem by framing the problem in the dual parameters \cite{csato2002sparse} of the GP that has previously been used for non-conjugate likelihood models in GPs \cite{adam2021dual}.

% CL section
%\textbf{Function-space methods for sequential learning} The central problem in continual learning is how to deal with the non-stationary nature of the training distribution, which causes the training process to overwrite the previously learnt parameters---leading to the model forgetting the previously learnt functions. 
%The approaches proposed in the CL literature can be categorized into inference-based, rehearsal-based, and model-based methods; we refer the reader to \citep{parisi2019continual, de2021continual} for complete reviews. The methods in the first category usually tackle this problem with weight-space regularization, such as EWC~\citep{kirkpatrick2017overcoming}, SI~\citep{zenke17a}), or VCL~\citep{nguyen-tuongModel2009} that induce retention of previously learned information in the weights alone.
%However, encoding functional knowledge in the weights does not guarantee a good quality of the predictions, thus a better to introduce a function-space regularization \citep{li2018lwf, benjamin2018measuring, titsias2019functional, buzzega2020dark, pan2020continual, rudner2022continual}. These methods bridge the gap between objective and rehearsal-based methods since they rely on a subset of the training data for each task to compute the regularization term. Recently proposed methods, such as DER~\citep{buzzega2020dark}, FROMP~\citep{pan2020continual}, and S-FSVI~\citep{rudner2022continual} achieve state-of-the-art performances among the objective-based techniques in several CL benchmarks.

\textbf{Function-space methods for sequential learning} 
Continual learning (CL) hinges on how to deal with the non-stationary training distribution, wherein the training process may overwrite previously learned parameters, causing catastrophic forgetting \citep{mccloskey1989catastrophic}. CL approaches fall into inference-based, rehearsal-based, and model-based methods (see \cite{parisi2019continual, de2021continual} for an overview). The methods in the first category usually tackle this problem with weight-space regularization, such as EWC~\citep{kirkpatrick2017overcoming}, SI~\citep{zenke17a}), or VCL~\citep{nguyen-tuongModel2009} that induce retention of previously learned information in the weights alone. However, encoding functional knowledge in the weights does not guarantee good quality of the predictions. Function-space regularization techniques~\cite{li2018lwf, benjamin2018measuring, titsias2019functional, buzzega2020dark, pan2020continual, rudner2022continual} address this by relying on training data subsets for each task to compute the regularization term. These methods bridge the gap between objective and rehearsal-based methods since they use a subset of the training data for each task to compute the regularization term. Recently proposed methods, \eg, DER~\citep{buzzega2020dark}, FROMP~\citep{pan2020continual}, and S-FSVI~\citep{rudner2022continual} achieve state-of-the-art performance among the objective-based techniques in several CL benchmarks. \looseness-2

\textbf{Uncertainty quantification in RL}
A key challenge in RL is balancing the trade-off between exploration and exploitation \cite{sutton2018reinforcement}.
That is, should an agent select actions that it knows will lead to high reward (exploitation), or should it
select new ones in hope to discover actions leading to higher reward (exploration).
One promising direction to balancing this trade-off is to model the uncertainty associated with a learned transition dynamics model and use it to guide exploration.
Prior work has taken an expectation over the dynamic model's posterior \cite{deisenrothPILCO2011,kamtheDataEfficient2018,chuaDeepReinforcementLearning2018},
sampled from it (akin to Thompson sampling but referred to as posterior sampling RL)
\todo{cite thopmson sampling paper dearden et al 1999}
\cite{osbandMoreEfficientReinforcement2013},
and used it to implement optimism in the face of uncertainty using upper confidence bounds \cite{curiEfficient2020}.
\todo{cite UCRL paper}
No single strategy has emerged as a go to method and we highlight that these strategies are only as good as their uncertainty estimates.
Previous approaches have used GPs \cite{deisenrothPILCO2011,kamtheDataEfficient2018},
ensembles of neural networks \cite{curiEfficient2020,chuaDeepReinforcementLearning2018}
and variational inference \cite{galImproving2016,houthooftVIME2017}.
%However, there has been little leveraging

% Another classical and provable exploration theory in RL is based on optimism
% \todo{cite UCRL paper}
% Recent work has even extended these ideas to deep RL \cite{curiEfficient2020}.

% An alternative approach is to sample from the dynamic model's posterior, akin to Thompson sampling but referred to as posterior sampling RL
% Another classical and provable exploration theory in RL is based on optimism in the face of uncertainty.
% \todo{cite UCRL paper}
% Recent work has even extended these ideas to deep RL \cite{curiEfficient2020}.
% These strategies are only as good as their uncertainty estimates.
% Prior work has learned dynamics models using GPs \cite{deisenrothPILCO2011,kamtheDataEfficient2018},
% ensembles of neural networks \cite{curiEfficient2020,chuaDeepReinforcementLearning2018}
% and variational inference \cite{galImproving2016,houthooftVIME2017}.

%%

% It should be written clearly that our method actually is doing a "functional" reharsal, because we store a set of inducing points that are replayed in order to check if the logits change.

% Other methods that could be worth citing are 
% Learning without forgetting \citep{li2018lwf} :  first paper proposing to use functional regularization computes a smoothed version of the current logits for the new examples at the beginning of each task, minimizing their drift during training. It doesn't have any set of inducing points
% \citep{benjamin2018measuring} this is not a proper CL paper, but it's usually referred in the functional regularization papers, because shows how to improve Adam with measuring the network difference in function space. They also have some easy CL examples there.

%Why we need uncertainty and adaptiveness
%How to adapt to new information is a limitation of current machine learning models. The problem is specifically relevant when we combine a model with a decision-making process. In such a sequential setting, an agent typically makes decisions in an environment and obtains new information and would ideally incorporate the information into their parameters. Not only do we need models that adapt to new informatoion but all express uncertainty over the models predictions. The reason being twofold: in many applications, \eg\ healthcare, autonomous driving is an essential requirement, and many advanced exploration techniques require uncertainty to determine where to query next.

\begin{figure}[t!]
  \centering
  % Set figure size
  \setlength{\figurewidth}{.31\textwidth}
  \setlength{\figureheight}{\figurewidth}
  %
  % Colours
  \definecolor{C0}{HTML}{DF6679}
  \definecolor{C1}{HTML}{69A9CE}
  %
  \begin{tikzpicture}[outer sep=0,inner sep=0]

    \newcommand{\addfig}[2]{
    \begin{scope}
      \clip[rounded corners=3pt] ($(#1)+(-.5\figurewidth,-.5\figureheight)$) rectangle ++(\figurewidth,\figureheight);
      \node (#2) at (#1) {\includegraphics[width=1.05\figurewidth]{./fig/#2}};
    \end{scope}
    %\draw[rounded corners=3pt,line width=1.2pt,black!60] ($(#1)+(-.5\figurewidth,-.5\figureheight)$) rectangle ++(\figurewidth,\figureheight);
    }

    % The neural network
    \addfig{0,0}{banana-nn}

    % The nn2svgp
    \addfig{1.1\figurewidth,0}{banana-nn2svgp}

    % The update
    \addfig{2.2\figurewidth,0}{banana-hmc}

    % The arrow
    \tikzstyle{myarrow} = [draw=black!80, single arrow, minimum height=14mm, minimum width=2pt, single arrow head extend=4pt, fill=black!80, anchor=center, rotate=0, inner sep=5pt, rounded corners=1pt]
    \tikzstyle{myblock} = [draw=black!80, minimum height=4mm, minimum width=7mm, fill=black!80, anchor=center, rotate=0, inner sep=5pt, rounded corners=1pt]
    \node[myarrow] (first-arr) at ($(banana-nn)!0.5!(banana-nn2svgp)$) {};
    \node[myblock] (second-arr) at ($(banana-nn2svgp)!0.5!(banana-hmc)$) {};

    % Arrow labels
    \node[font=\scriptsize\sc,color=white] at (first-arr) {\our};
    \node[font=\scriptsize\sc,color=white] at (second-arr) {\normalsize$\bm\approx$};

    % Labels
    \node[anchor=north, font=\small] at ($(banana-nn) + (0,-.55\figureheight)$) {Neural network prediction};
    \node[anchor=north, font=\small] at ($(banana-nn2svgp) + (0,-.55\figureheight)$) {Sparse function-space representation};
    \node[anchor=north, font=\small] at ($(banana-hmc) + (0,-.55\figureheight)$) {HMC result as baseline};

  \end{tikzpicture}
  \newcommand{\mycircle}{\protect\tikz[baseline=-.6ex]\protect\node[circle,inner sep=2pt,draw=black,fill=C0,opacity=.5]{};}
  \newcommand{\mysquare}{\protect\tikz[baseline=-.6ex]\protect\node[inner sep=2.5pt,draw=black,fill=C1,opacity=.5]{};}
  \newcommand{\myinducing}{\protect\tikz[baseline=-.7ex]\protect\node[circle,inner sep=1.5pt,draw=black,fill=black]{};}
  %
  \caption{\textbf{Uncertainty quantification} for binary classification (\mysquare~vs.~\mycircle). We convert the trained neural network (left) to a sparse GP model that summarizes all data onto a sparse set of inducing points~\myinducing\ (middle). This gives similar behaviour as would running full Hamiltonian Monte Carlo (HMC) on the original neural network model weights (right). Marginal uncertainty depicted by colour intensity.\looseness-1}
  \label{fig:banana}  
\end{figure}

\section{Background: Function-space representation of neural networks}
\label{sec:methods}
%
%In this section, we recap how a trained neural network (NN) can be converted to a Gaussian process by locally linearising its weights.
% a GP posterior -- can be obtained
% around the Maximum a Posteriori (MAP) weights.

% by linearising a neural network around the Maximum a Posteriori (MAP) weights.
% a GP posterior -- can be obtained
%
In supervised learning, given a data set $\dataset = \{(\vx_{i} , \vy_{i})\}_{i=1}^{N}$, with input $\vx_i \in \inputDomain$ and output $\vy_i \in \outputDomain$ pairs, the weights $\weights \in \R^{P}$ of a neural network, $f_\mathbf{w} : \inputDomain \to \outputDomain$, are usually trained to minimize the (regularized) empirical risk,
%
\begin{equation} \label{eq-empirical-risk}
  \weights^{*} = 
  \arg \min_{\weights} \mathcal{L}(\dataset,\weights) =
  \arg \min_{\weights} \textstyle\sum_{i=1}^{N} \ell(f_\weights(\mathbf{x}_{i}), \mathbf{y}_i) + \delta \mathcal{R}(\weights).
\end{equation}
%
If $\ell(f_\weights(\vx_{i}), \vy_i) = -\log(p(\vy \mid f_\weights(\vx_{i}))$ and the regularizer $\mathcal{R}(\weights) = \frac{1}{2}\|\weights\|^{2}_2$, then we can view \cref{eq-empirical-risk} as the maximum {\it a~posteriori} (MAP) solution to a Bayesian objective, where the regularization weight takes the role of a prior precision parameter, \ie, $p(\vw) = \Norm(\vzeros, \delta^{-1} \MI)$.
%Bayesian inference offers a principled approach to quantifying uncertainty in neural networks.
%The goal is to find the posterior over the weights ${p(\vw \mid \vy) \propto p(\vy \mid f_{\weights}(\vx)) \, p(\weights)}$ as it represents our belief in the parameters after combining data $\dataset$ with our prior $p(\vw)$.
%Although the true posterior $p(\vw \mid \dataset)$ is intractable given the non-linearities of the neural network, one can resort to sampling techniques such as Hamiltonian Monte-Carlo (HMC).
%However, for most applications their high computational costs make them impractical so we need approximate inference techniques.
%In \cref{fig:banana}, we show a qualitative example of the induced function-space posterior obtained with our method compared to the posterior obtained through HMC sampling on the Banana toy dataset. 
The posterior over the weights ${p(\vw \mid \dataset) \propto p(\vy \mid f_{\weights}(\vx)) \, p(\weights)}$ is generally not available in closed form. Sampling methods that characterize the posterior with a finite set of samples in weight-space---such as the Hamiltonian Monte Carlo (HMC) baseline used in \cref{fig:banana} (right)---are general-purpose, but comuptationally heavy and impractical for downstream applications.

\textbf{Function space}
%As a neural network is a deterministic mapping, the weight-space posterior induces a distribution over the function values.
%Intuitively, if one was to sample from the weight posterior, the corresponding functions created can
%be viewed as perturbed versions of the function at the MAP estimate $f_{\vw^*}$.
%In most applications, we care about predictions from the neural network and not the weights themselves.
%As such, it is the distribution over function values that we are actually interested in.
While neural networks are deterministic mappings defined by their weights, the ultimate goal of training a neural network is to optimize the function it represents, not the weights themselves. Consequently, the weight-space posterior corresponds to a distribution over function values. Imagine sampling from the weight posterior; the resulting functions could be viewed as perturbations of the function at the MAP estimate $f_{\vw^*}$. This perspective aligns better with the main objective in most applications: making representative predictions given the observations we have.

\textbf{Linearization gives rise to a Gaussian process}
Our goal is to capture the distribution over the neural network model functions through their first two moments. The first moments characterize a Gaussian process with a mean function $\mu(\cdot)$ and a covariance function (kernel) $\kappa(\cdot,\cdot)$. 
Recent work \cite{khan2019approximate,maddox2021fast} has shown that linearizing approximations in the weight space lead to function-space equivalent approximations.
As Gaussian distributions remain tractable under linear transformations, a linear function in terms of parameters can be converted from the weight space to the function space (see Ch.~2.1 in \cite{rasmussen2006gaussian}) as follows:
%
\begin{equation} \label{eq:weight_func}
f_\weights(\vx) \approx 
%g_\weights(\mathbf{x}) = 
\phi^\top\!(\vx) \, \vw \quad\implies\quad \mu(\vx) = 0 \quad \text{and} \quad \kappa(\vx, \vx') = \frac{1}{\delta} \phi^\T\!(\vx) \, \phi(\vx').
\end{equation}
% The posterior structure directly relates to the optimization loss around the MAP weights $\vw^*$.
A common approach is to approximate the correlation structure of a distribution centred at the MAP estimate as done in the Laplace-GGN~\cite{khan2019approximate, daxberger2021laplace, maddox2021fast}. 
The Laplace-GGN takes the MAP solution and approximates the Hessian of the loss function
with the GGN.
Following the work of \citet{khan2019approximate}, we can view this approximation as building an approximate linear model of the neural network as $f_{\weights^*}(\vx) \approx 
%g_{\weights}(\vx) = 
\Jac{\weights_*}{\vx} \, \weights$, where $\Jac{\weights}{\vx} \coloneqq \left[ \nabla_\weights f_\weights(\vx)\right]^\top \in \R^{C \times P}$ is the Jacobian at the MAP.
Using the Hessian of the approximate model, we arrive at the Laplace-GGN approximate posterior for $q(\vw)$. 
Therefore, the Laplace-GGN linear model can be used to convert our weight space model to the function space by,
\begin{equation} 
\label{eq-laplace-approx-function-space} 
% g(\vx) \sim \GP \left( \mu(\vx), \kappa(\vx, \vx') \right) \quad \text{with} \quad
  \mu(\vx) =  0 \quad \text{and} \quad
  \kappa(\vx, \vx')
  = \frac{1}{\delta} \Jac{\weights^*}{\vx} \, \JacT{\weights^*}{\vx'}, 
\end{equation}
where the kernel is the so-called Neural Tangent Kernel (NTK, \cite{jacot2018neural}). In \cref{eq-laplace-approx-function-space}, we restrict the function output to a single dimension for notational simplicity, yet this formulation can be readily extended to the multi-output scenario. With this kernel function, we can combine it with the data set $\dataset$ to construct the posterior $q(\vf)$. 
%
\citet{khan2019approximate} took a similar approach, but instead of fitting the GP posterior to the actual $\vy$, they rely on a transformation of $\vy$ referred as {\em pseudo}-data. Similarly, \citet{immer2021improving} obtain the same covariance function but use a different mean because they rely on a first-order approximation of the neural network to form a Bayesian GLM model. Both approaches attempted to adjust the GP posterior mean function to ensure the predictions are from the neural network $f_{\vw^*}(\vx)$. 
Contrarily, our methodology intends to make predictions directly from the derived GP, thus evading such adjustments. Next, we will present the general formulation of the sparse functional representation of our trained neural network based on the non-sparse approximation.

\section{\our: Sparse function-space represenation of neural networks}
%\paragraph{GP in the dual parameters}
The seminal work by \citet{csato2002sparse} (parameterization Lemma~1) gives a parameterization for the posterior process $f_{\textrm{post}}$ that can be found through the Bayesian update using the GP prior and the likelihood function. This gives a `dual' parameterization, $\valpha$ and $\vbeta$,
%
\begin{equation}  \label{eq:gp_pred}
  \myexpect_{p(f_i \mid\vy)}[f_i]= \vk_{\vx i}^\top \valpha \quad \text{and} \quad
  \mathrm{Var}_{p(f_i \mid \vy)}[f_i] = k_{ii} - \vk_{\vx i}^\top ( \MKxx + \diag(\vbeta)^{-1})^{-1} \vk_{\vx i},
\end{equation}
%
where the $ij$\textsuperscript{th} entry of the matrix $\MKxx \in \R^{N \times N}$ is $\kappa(\vx_i,\vx_j)$, $\vk_{\vx i}$ denotes a vector where each $j$\textsuperscript{th} element is $\kappa(\vx_i, \vx_j)$, and $k_{ii} = \kappa(\vx_i, \vx_i)$.  The dual parameters are the vectors $\valpha, \vbeta \in \R^{N}$, defined as:\looseness-1
%
\begin{equation}
  \label{eq:dual_param}
  \alpha_i \coloneqq \myexpect_{p(\vw \mid \vy)}[\nabla_{f_i}\log p(y_i \mid f_i)]
  \quad \text{and} \quad
  \beta_i \coloneqq - \myexpect_{p(\vw \mid \vy)}[\nabla^2_{f_i f_i}\log p(y_i \mid f_i)] .
\end{equation}
%
The relationships specified above are valid for generic likelihoods, and involve no approximations since the expectation is under the exact posterior. This higlights the fact that our approximate inference technique that will follow, while usually viewed as a posterior approximation, can be alternatively interpreted as an approximation of the expectation of gradients of loss/likelihoods.

Originally \citet{csato2002sparse} iteratively find dual variables using an expectation propogation (EP, \cite{minka2001expectation}) method. More recently \cite{khan2017conjugate,adam2021dual}, has shown this relationship for variational Gaussian processes where the above expectation is with respect to the approximate variational posterior. The dual parameter $\valpha$ appears also in non-Bayesian approaches, \eg, support vector machines \citep{cortes1995support}, whose origins are found in work by \citet{kimeldorf1971some}. While in \citet{wilkinson2023bayes}, they show links between linearization methods and how they solve the \cref{eq:dual_param}.

\paragraph{Dual parameters from NN}
Given that we use a Laplace approximation of the neural network, we remove the expectation over the posterior (see Ch.~3.4.1 in \cite{rasmussen2006gaussian} for derivation) and we get the the following formulation of the dual variables,
%
\begin{equation}
  \label{eq:dual_param_laplace}
  \hat{\alpha}_i \coloneqq \nabla_{f}\log p(y_i \mid f) |_{f=f_i}
  \quad \text{and} \quad
  \hat{\beta}_i \coloneqq - \nabla^2_{f,f}\log p(y_i \mid f) |_{f=f_i}.
\end{equation}
%
Substiuting \cref{eq:dual_param_laplace} into \cref{eq:gp_pred}, we obtain our GP model based on the converged neural network. Again, this is similar to what was derived in \citet{immer2021improving} for the posterior variance function. However, they use the $f_{\vw^*}$ for the posterior mean and do not emphasize the significance of the dual variables. The problem with \cref{eq:gp_pred} is that to make predictions and compute variances we must incur a cost of $O(N^3)$ which limits the use of the GP on large data sets.

\paragraph{Sparsifying the NNGP}
\label{sec:sparse-dual-gp}
%
Sparse Gaussian processes reduce the computational complexity by representing the GP as a low-rank approximation induced by a sparse set of inputs points (see \cite{quinonero2005unifying} for an early overview). Given that we have computed the dual variables derived from our neural network predictions and a kernel function, we could essentially employ any of these sparsification methods. In this work, we opt for the approach suggested by \citet{titsias2009variational} (also used in the DTC approximation, see \cite{quinonero2005unifying}), which defines the marginal predictive distribution as $q_{\vu}(f_i)  = \int p(f_i  \mid \vu) \, q(\vu) \, \mathrm{d}\vu$. Given that we have $p(f_i \mid \vu)$ determined by our GP prior, the goal is to find a $q(\vu)$.

As demonstrated in \cite{adam2021dual}, the posterior under this model bears a structure akin to \cref{eq:gp_pred}. The authors of that paper see the significance of the dual variables for approximate inference, but write them using the natural parameterization, primarily because this form is more suitable for optimization through natural gradients. In order to link the dual variables defined in \cref{eq:dual_param}, we write them in the sparse GP using the dual variables
%
\begin{equation} \textstyle
  \valpha_{\vu}  =  \sum_{i=1}^N  \vkzi \, \hat{\alpha}_{i}
  \quad \text{and} \quad
  \MBeta_{\vu} =  \sum_{i=1}^N \vkzi \,\hat{\beta}_{i} \, \vkzi^{\T} ,    
\label{eq:dual_sparse}
\end{equation}
%
where the sparse dual variables are now a sum over \emph{all data points}, with $\valpha_{\vu} \in \R^{M}$ and $\MBeta_{\vu} \in \R^{M  \times M}$. Using this sparse definition of the dual variables, our sparse GP posterior takes the following form:
\begin{equation}\label{eq:dual_sparse_post}
   \myexpect_{q_{\vu}(\vf)}[f_i] = \vkzs^{\T} \MKzz^{-1} \valpha_{\vu}
   \quad \text{and} \quad 
   \textrm{Var}_{q_{\vu}(\vf)}[f_i]  = k_{ii} - \vkzs^\top [\MKzz^{-1} - (\MKzz + \MBeta_{\vu})^{-1} ]\vkzs
\end{equation}
where $\MKzz$ and $\vkzs$ are defined similarly to $\MKxx$ and $\vk_{\vx i}$ but over the inducing points $\{\vz_j\}_{i=1}^M$, $\vz \in \R^{D}$, instead of the full data set $\dataset$. The key quantities we need to make predictions from our sparse GP from the converged neural network are $(\hat{\alpha}_i, \hat{\beta}_i)$ \cref{eq:dual_param_laplace} and a kernel function $\kappa$ . Contrasting \cref{eq:dual_sparse_post} and \cref{eq:gp_pred}, we can see that the computational complexity went from $\mathcal{O}(N^3)$ to $\mathcal{O}(M^3)$, with $M \ll N$.  Crucially, given the structure of our probabilistic model, our sparse dual variables \cref{eq:dual_sparse} are a compact representation of the full model projected using the kernel.

Unlike our approach, \citet{immer2021improving} employ a subset of the data points to construct a sparse GP model, reducing computational complexity. However, this method could be considered less principled for building a sparse model, as it ignores contribution from data. It is important to note that the sparsifying process is independent of the approximate inference technique because the dual variables of \cref{eq:dual_param} are computed using \cref{eq-empirical-risk} and a Laplace approximation.
More complicated inference techniques, such as variational inference, could be used, but given its simplicity, in our experiments we used the Laplace-GGN approximation with the trained neural network. Furthermore, our dual variable view of approximate inference means we do not need to retrain a separate sparse GP model.%\todo{shouldn't this sentence be clear from the context? we derive the mean/kernel function that define the process, then we have a sparse GP already}

\section{Maintaining a sparse function-space representation in sequential learning}
\label{sec:sequential}
%
We have presented a method for converting any trained neural network into a sparse function-space representation. The sparse representation opens an avenue for maintaining (and updating) a representation of the neural network, an important feature for sequential learning, such as continual learning.\looseness-1
%In this section, we demonstrate how sequential learning methods can benefit from our function-space representation and it's associated uncertainty estimates.

\paragraph{\our for continual learning}
% Definition of CL -> a sequence of tasks

In the continual learning setting, the training is divided into $T$ tasks, each with its training data set $\dataset_t = \{(\mathbf{x}_{i}, \mathbf{y}_{i})\}_{i=1}^{N_t}$, which after the task is discarded and cannot be accessed in the following stages of the training. Rehearsal and function regularization-based models keep a subset of the training data for each task to help the model alleviate forgetting. In the same way, after each task, we can use our \our method to build a compact representation of the neural network and then exploit this representation to build a functional regularizer to add to the training loss for the subsequent tasks. Recently proposed GP-based regularizers \cite{ pan2020continual, rudner2022continual} have shown better performances than weight space equivalents getting state-of-the-art on CL benchmarks. However, they are limited to use subset approximations of the posterior \cref{app:cl-experiment}..
%, which in practice means randomly selecting some data computing its posterior at only those points and then using that as the regularizer. 
A more principled approach would be to sparsify the resultant GP into a low-rank approximation to ensure information coming from the entire task dataset is not lost. %, essentially what the sparse dual variables do.   \todo{this term is replacing the missing data}
% Where the regularizer comes from
At the end of the training for each task $t$, we compute the Laplace-GGN approximated kernel function for the current model and randomly select $M$ inducing points. Given the issues described in the previous sections, related to computing the full covariance posterior $\MKxx$ for $\dataset_t$,\todo{this should be $\kappa(\MX_t, \MX_t)$} we can exploit the dual parameterization to efficiently encode the information from  $\dataset_t$ using the dual parameterization, as follows, 
%
\begin{equation}\textstyle
 	\bar{\MB}^{-1}_t = \MKzz^{-1} \MBeta_\vu \MKzz^{-1} \in \R^{M \times M}, 
 	\quad \text{s.t.} \quad
 	\MBeta_{\vu} =  \sum_{i \in \dataset_t} \vkzi \,\hat{\beta}_{i} \, \vkzi^{\T} ,    
 	% \quad \forall t \in [1, T]
\end{equation}
%
where $\MKzz$ and $\vkzi$ are the gram matrix and vector computed on the inducing points $\MZ_t = \{\vz_i\}_{i=1}^M$ for task $t$ .\
%
We keep a memory buffer $\mathcal{M} = \{(\MZ_t, \vu_t, \bar{\MB}^{-1}_t)\}_{t=1}^T$, where $\MZ_t$ are the inducing points selected from $\dataset_t$ and  $\vu_t = f_{\vw_t^*}(\MZ_t)\in \R^{M}$ are the neural network outputs associated to them, computed with the MAP estimate $\weights_t^*$ after task $t$. Note that, for notational convenience, we are extending our neural network notation to be compatible with multiple inputs points, \ie, each $i$\textsuperscript{th} entry is $f_{\weights_t^*}(\vz_i)\in \R$. Moreover, we restrict ourselves to single-output neural networks and refer the reader to further details on the multi-output setting to \cref{sec:cl_multioutput}.

After each task, we update $\mathcal{M}$, and we can construct a regularizer term, which aims to reduce the drift of the current model from diverging on the inducing points $\MZ_s$ for all previous tasks $ s = 1, \ldots, t-1$. This means that for training  the successive task, the new objective to minimize becomes $\weights_t^* = \arg\min_{\weights} \mathcal{L}(\mathcal{D}_t, \weights) + \tau \, \mathcal{R}_\textrm{\our}(\weights, \mathcal{M})$, with the regularizer
%
\begin{equation}\label{eq-cl-regularizer}\textstyle
  \mathcal{R}_\mathrm{\our}(\weights, \mathcal{M}) = \frac{1}{2} \sum_{s=1}^{t-1} \frac{1}{M} 
	\left\lVert 
	f_{\weights}(\MZ_{s}) - \vu_t % f_{\weights_{s}}(\MZ_s)
	\right\rVert_{\bar{\MB}^{-1}_{s}}.
\end{equation}

%%The resulting regularizer can be computed as follows
%%\begin{equation}
%%	\mathcal{R}_\textit{SFR}(\weights, \mathcal{M}) = \sum_{s=1}^{t-1} \frac{1}{m} \left[\left(f_{\weights}(\MZ_{s}) - f_{\weights_{s}}(\MZ_s) \right)^\T \bar{\MB}^{-1}_{s} \left(f_{\weights}(\MZ_{s}) - f_{\weights_{s}}(\MZ_s) \right) \right].
%%\end{equation}
%%
%\begin{equation}
%	\mathcal{R}_\mathrm{\our}(\weights, \mathcal{M}) = \frac{1}{2} \sum_{s=1}^{t-1} \frac{1}{M} 
%	\left\lVert 
%	f_{\weights}(\MZ_{s}) - \vu_t % f_{\weights_{s}}(\MZ_s)
%	\right\rVert_{\bar{\MB}^{-1}_{s}}.
%\end{equation}
%%\todo{choose one option}
%% where $f_\vw(\MZ_{s}) \in \R^m$ are the outputs of neural networks with the current weights $\vw$ and stored function outputs of the neural network at the MAP optimum for task $s$, i.e., each $i$th entry is $f_\cdot(\vz_i) \in \R$.
%%\begin{equation}
%%	\argmin_{\weights} \sum_{i=1}^{n} \ell(f_\weights(\mathbf{x}_{i}), \mathbf{y}_i) + \delta \mathcal{R}(\weights) + \tau \mathcal{R}_\textit{SFR}(\weights)
%%\end{equation}
%Then, when the training on the successive task, the new objective to minimize takes the following form
%\begin{equation}
%	\weights_t^* = \argmin_{\weights} \mathcal{L}(\mathcal{D}_t, \weights) + \tau \mathcal{R}_\textit{SFR}(\weights, \mathcal{M})
%\end{equation}

\section{Experiments}
\label{sec:experiments}
%
We present a series of experiments designed to demonstrate the versatility and efficacy of our proposed \our framework. We seek to find answers to the following questions: 
\textbf{Predictions:}~Does our method's ability to consider the full data set offer benefits over subset function-space methods? (\cref{sec:uci}) How do predictions with our sparse function-space approximation compare to weight-space approximations? 
\textbf{Function-space updates:} How fast are our function-space updates relative to retraining from scratch? Do they improve predictive performance? Are they as good as retraining from scratch? \todo{what do we have on this?}
\textbf{Representation:} Can a sparse function-space representation help retain knowlegde from previous tasks in CL? (\cref{sec:cl-exp})
\textbf{Uncertainty:} Can our uncertainty estimates help exploration in RL under sparse rewards? (\cref{sec:rl-exp})

%Initially, we explore the capability of our method to capture predictive uncertainty through supervised learning tasks using UCI datasets. Subsequently, we extend our supervised learning experiment to image datasets, demonstrating our method's robustness and adaptability to more complex, high-dimensional data. Next, we investigate the potential of our method in a continual learning context, where we update the neural network representation in response to incoming data without necessitating retraining. Lastly, we delve into the realm of reinforcement learning to ascertain the applicability of our approach in an environment that requires sequential decision-making. Each experiment is designed to illuminate a different facet of our method, thereby exemplifying its broad range of applicability and potential for future utilization.

%
%Our experiments seek to answer the following questions:
%\begin{enumerate}
%  \item \textbf{Predictions:} How do predictions with our sparse function-space approximation compare to weight-space approximations? Does our method's ability to consider the full data set offer benefits over subset function-space methods?
%  % \item Does our method's ability to consider the full data set offer benefits over subset function-space methods?
%  \item \textbf{Function-space updates:} How fast are our function-space updates relative to retraining from scratch? Do they improve predictive performance? Are they as good as retraining from scratch?
%  \item \textbf{Uncertainty:} How good are our uncertainty estimates? Can they be used in downstream settings like RL?
%  \item \textbf{Representation:} Is our sparse function-space representation useful for continual learning?
%\end{enumerate}

\subsection{Capturing uncertainty in UCI tasks under supervised learning}
\label{sec:uci}
%

\begin{table}[t!] 
  \centering\scriptsize
  \caption{Negative log predictive density (NLPD) (lower better) for the proposed model TODO} 
	\label{tbl:uci}
	% Control table spacing
	\renewcommand{\arraystretch}{1.}
	\setlength{\tabcolsep}{2pt}
	\setlength{\tblw}{0.14\textwidth}

	% Custom error formatting
	\newcommand{\val}[2]{%
		$#1$\textcolor{gray}{\tiny ${\pm}#2$}
	}

    % THE TABLE NUMBER ARE GENERATED BY A SCRIPT	
	\input{tables/uci.tex}
\end{table}
We trained a two-layer MLP for each of the $8$ classification tasks in UCI, see \cref{tbl:uci} for a list of the datasets. The neural network training was done using MAP, with a prior $\mathcal{N}(0,\delta^{-1} \MI)$ with prior precision $\delta$. After training, we constructed the SVGP dual and used the resulting SVGP model for uncertainty quantification. Depending on task dimension and complexity, our model can match the NN MAP in NLPD performance even with a low number of inducing points, see \cref{tbl:uci}. We performed hyperparameter search over the prior precision, and ran the experiment over $10$ seeds. During the initial NN training, we used a learning rate of $1e-3$.

\subsection{Supervised learning on image data sets}

\begin{table}[t!] 
  \centering\scriptsize
  \caption{Metrics for supervised learning with image data. SVGP is our method, and the number inside parentheses is the number of inducing points. SVGP NN is a modification of our method, where the mean is from the NN directly and the variance is from the GP model} 
	\label{tbl:imagesuper}
	% Control table spacing
	\renewcommand{\arraystretch}{1.}
	\setlength{\tabcolsep}{6pt}
	\setlength{\tblw}{0.15\textwidth}

	% Custom error formatting
	\newcommand{\val}[2]{%
		$#1$\textcolor{gray}{\tiny ${\pm}#2$}
	}

    % THE TABLE CAN BE FORMATTED MORE LIKE THIS
    \begin{tabular}{l l C{\tblw} C{\tblw} C{\tblw} C{\tblw}}
    \toprule
    & Method & ACC~$\uparrow$ & NLPD~$\downarrow$ & ECE~$\downarrow$ & OOD-AUC~$\uparrow$  \\
    \midrule
    \multirow{2}{*}{FMNIST} 
    & MAP & \val{0.000}{0.000} & \val{0.000}{0.000} & \val{0.000}{0.000} & \val{0.000}{0.000} \\
    & BNN predictive & \val{0.000}{0.000} & \val{0.000}{0.000} & \val{0.000}{0.000} & \val{0.000}{0.000} \\
    & BNN predictive \cite{todo} & \val{0.000}{0.000} & \val{0.000}{0.000} & \val{0.000}{0.000} & \val{0.000}{0.000} \\
    & GLM predictive & \val{0.000}{0.000} & \val{0.000}{0.000} & \val{0.000}{0.000} & \val{0.000}{0.000} \\
    & GP predictive & \val{0.000}{0.000} & \val{0.000}{0.000} & \val{0.000}{0.000} & \val{0.000}{0.000} \\
    & \our (NN) & \val{0.000}{0.000} & \val{0.000}{0.000} & \val{0.000}{0.000} & \val{0.000}{0.000} \\
    & \our & \val{0.000}{0.000} & \val{0.000}{0.000} & \val{0.000}{0.000} & \val{0.000}{0.000} \\    
    \midrule
    \multirow{2}{*}{CIFAR-10} 
    & MAP & \val{0.000}{0.000} & \val{0.000}{0.000} & \val{0.000}{0.000} & \val{0.000}{0.000} \\
    & BNN predictive & \val{0.000}{0.000} & \val{0.000}{0.000} & \val{0.000}{0.000} & \val{0.000}{0.000} \\
    & BNN predictive \cite{todo} & \val{0.000}{0.000} & \val{0.000}{0.000} & \val{0.000}{0.000} & \val{0.000}{0.000} \\
    & GLM predictive & \val{0.000}{0.000} & \val{0.000}{0.000} & \val{0.000}{0.000} & \val{0.000}{0.000} \\
    & GP predictive & \val{0.000}{0.000} & \val{0.000}{0.000} & \val{0.000}{0.000} & \val{0.000}{0.000} \\
    & \our (NN) & \val{0.000}{0.000} & \val{0.000}{0.000} & \val{0.000}{0.000} & \val{0.000}{0.000} \\
    & \our & \val{0.000}{0.000} & \val{0.000}{0.000} & \val{0.000}{0.000} & \val{0.000}{0.000} \\    
    \bottomrule
    \end{tabular}

    % THE TABLE NUMBER ARE GENERATED BY A SCRIPT	
	%\input{tables/img_super.tex}
\end{table}

Similarly to the UCI experiments, we trained neural networks (CNN, architexture matches \citet{immer2021improving}) for the FMNIST and CIFAR10 classification tasks. After training, we computed the duals for NN2SVGP and used the resulting SVGP for uncertainty estimates, see \cref{tbl:imagesuper}. We performed hyperparameter optimization over the prior precision, and found the optimal value to be [insert]. We ran the experiment over $5$ seeds. During the NN training, we used a batch size of $512$ and an Adam optimizer with learning rate $1e-3$.

In addition to uncertainty estimates, our method can update the posterior when given new data. We gave the SVGP model 10 \% of the test dataset used for evaluation, and compare the performance of the SVGP model and the retrained NN in [INSERT TABLE].

\subsection{Updating the  network representation in continual learning}
%% Very rough draft
This experiment's section is meant to demonstrate the quality of our sparse representation to retain previous knowledge in the Continual learning setting. Unfortunately the CL literature is based on methods that require very diverse set of hyperparameters 
% Idea: it is very difficult to have everything under control, because model changes, settings change, optimizer change, and so on. Then what we did was trying to replicate as close as possible the results for the competing methods by using their codebases where available and choosing the hyperpameters accordingly to their choices.

% Why SH?
% TODO: there are at least two papers that clearly say MH is unrealistic and too simple as a consequence is not a good way to assess the quality of the method

% Architectures
We conduct the experiments on a set of three CL benchmarks Sequential-MNIST (S-MNIST), Sequential-FashionMNIST (S-FMNIST), and Permuted-MNIST (P-MNIST).  Following \cite{rudner2022continual, pan2020continual}, we used a two-layer MLP with 256 hidden units with ReLU activation for S-MNIST and S-FMNIST, and a two-layer with 100 hidden units for P-MNIST.

% Methods
We compared our method to 1) weight-regularization based methods EWC, SI and VCL (which also has the addition of coresets); 2) function-based regularization based methods DER, FROMP, S-FSVI, SFR; 3) an ablation for which we replace $\bar{\MB}^{-1}$ with an identity matrix $\MI_m$, \ie which is equivalent to assuming a GP with iid realization \todo{Not sure this is 100$\%$ correct}.
% Inducing points
Following \citep{rudner2022continual}, we use 200 inducing points per tasks % this one can be difficult to justify since S-FSVI samples 200 context points and then iteratively samples 40 coresets points (or at least that's what I got)
on all the inducing points, and a lower number of inducing point for S-MNIST, to investigate if our method was still capable of capture more information due to its sparse formulation, as explained in \cref{todo}.

%% Questions: Should we also mention the Identity outside when not trained on those classes or just leave that for the appendix?

%% Comments on the results:
We get the best results on S-MNIST both 40 and 200 pts. per task

On S-FMNIST we are on par with DER, which is however exploiting a reservoir buffer technique, that iteratively samples new "inducing points"

On P-MNIST we are lower than S-FSVI, however % I'm still running experiments, so we can hope that it is not the best overall in the end

\begin{table}[t!] 
  \centering\scriptsize
  \caption{
  TODO: CL Experiments. $^*$ Methods relying only on weight regularization. 
  \color{blue}{Results from \cite{rudner2022continual}.} 
  %\color{red}{Preliminary results need more runs or hyperparam tuning}
  }
	\label{tbl:cl_table_1}
	% Control table spacing
	\renewcommand{\arraystretch}{1.}
	\setlength{\tabcolsep}{2pt}
	\setlength{\tblw}{0.14\textwidth}

	% Custom error formatting
	\newcommand{\val}[2]{%
		$#1$\textcolor{gray}{\tiny ${\pm}#2$}
	}

	\input{tables/cl_table_1}
\end{table}

\subsection{Reinforcement learning under sparse rewards}
\label{sec:rl-exp}
% We now demonstrate the quaility of \our's uncertainty estimates by showing that it can improve sample efficiency in RL.
% Balancing the trade-off between exploration and exploitation is a key challenge in RL \cite{sutton2018reinforcement}.
% Should an agent select actions that it knows will lead to high reward (exploitation), or actions which it has not select before in the
% hope to discover new high reward actions (exploration)?

% when used in conjuction with an
% uncertainty-guided exploration strategy.

% In this section, we show that \our's uncertainty estimates can be used to balance this trade-off in a model-based RL algorithim,
% demonstrating the quality of our uncertainty estimates.

% A key challenge in RL is balancing the trade-off between exploration and exploitation.
% That is, should an agent select actions that it knows will lead to high reward (exploitation), or should it
% select new ones in hope to discover actions leading to higher reward (exploration).
% One promising direction is to model the uncertainty associated with a learned transition dynamics and use it to guide exploration.
% Prior work has learned dynamics models using GPs \cite{deisenrothPILCO2011,kamtheDataEfficient2018},
% ensembles of neural networks \cite{curiEfficient2020,chuaDeepReinforcementLearning2018}
% and variational inference \cite{galImproving2016,houthooftVIME2017}.
% There are then many ways to leverage uncertainty in model-based RL, for example,
% taking an expectation over epistemic uncertainty  \cite{deisenrothPILCO2011,kamtheDataEfficient2018,chuaDeepReinforcementLearning2018},
% sampling from the posterior, akin to Thompson sampling but referred to as posterior sampling RL
% \cite{osbandMoreEfficientReinforcement2013},
% and methods based on upper confidence bounds (UCB) \cite{curiEfficient2020}.

In this section, we show that \our's uncertainty estimates can be used to guide exploration in model-based RL;
demonstrating the quality of our uncertainty estimates.
% Uncertainty-guided decision making strategies
We test \our by using it to learn a dynamics model within a model-based RL strategy which uses posterior sampling to guide exploration
\cite{osbandMoreEfficientReinforcement2013,osbandWhyPosteriorSampling2017}.
We use the cart pole swing up task in MuJoCo \cite{todorov2012mujoco}, a classic benchmark for nonlinear control, see \cref{fig:rl}.
The goal is to swing the pole up and balance it around the upward position.
We increase the difficulty of exploration by using a sparse reward function.
We refer the reader to \cref{app:rl} for an overview of the reinforcement learning problem, details of the algorithim we implement and our experiment's configuration.

\cref{fig:rl} shows training curves using \our as the dynamic model (cyan) as well as a Laplace-GGN with GLM predictions (red), an ensemble of neural
networks (green) and a basic MLP wih no uncertainty (magenta).
To ensure a fair comparison, we use the same MLP architecture/training scheme and use them in the same model-based RL algorithm, detailed in \cref{app:rl}.
% that \our's uncertainty estimates are useful for guiding exploration as it converges in fewer episodes.
We also compare to deep deterministic policy gradient (DDPG) \cite{lillicrapContinuousControlDeep2016}, a model-free RL algorithim (yellow).
The training curves show that \our's uncertainty estimates are useful for guiding exploration as it converges in fewer episodes, i.e.
it is more sample efficient.
As expected, the MLP strategy (without uncertainty) was not able to successfuly explore the environment.

% We compare \our (cyan) to a Laplace-GGN with GLM predictions (red), an ensemble of neural networks (green) and a basic MLP wih no uncertainty (magenta).
% To ensure a fair comparison, we use the same MLP architecture/training scheme and use them in the same model-based RL algorithm, detailed in \cref{app:rl}.
% See \cref{app:rl} for more details of our experiments.
% As expected DDPG is the least sample inefficient
% The training curves in \cref{fig:rl} show that \our's uncertainty estimates are useful for guiding exploration as it converges in fewer episodes.

% \cite{deisenrothPILCO2011},
% Posterior sampling

% based on

% Model-based RL algorithims are more sample efficient than their model-free counterparts as they learn a model of the transition dynamics and
% use it to augment the RL loop.

% Prior work has used show that modelling uncertainty in the transition dynamics can further improve exploration and thus sample efficiency.
% For example, with GPs \cite{deisenrothPILCO2011,kamtheDataEfficient2018},
% ensembles of neural networks \cite{curiEfficient2020,chuaDeepReinforcementLearning2018}
% and variational inference \cite{galImproving2016,houthooftVIME2017}

% We now demonstrate the quaility of \our's uncertainty estimates by showing that it can improve sample efficiency in RL.

% In this section, we show that \our's uncertainty estimates can be used to balance this trade-off in a model-based RL algorithim,
% demonstrating the quality of our uncertainty estimates.
% In this section, we demonstrate  that our method's principled uncertainty estimates can be used to balance this trade-off in a model-based RL algorithim.

% It is an inherently unstable and underactuated mechanical system.

% by using it to learn a dynamics model and
% evaluate the
% We test \our by using it to learn a dynamics model
% our method on the cart pole swing up task in MuJoCo \cite{todorov2012mujoco}, a classic benchmark for nonlinear control, see \cref{fig:rl}.
% The goal is to swing the pole up and balance it around the upward position.
% We increase the  difficulty of exploration in this environment by using a sparse reward function.

% We compare our method to a Laplace-GGN with GLM predictions (red), an ensemble of neural networks (green) and a basic MLP wih no uncertainty (magenta).
% To ensure a fair comparison, we use the same MLP architecture/training scheme and use them in the same model-based RL algorithm, detailed in \cref{app:rl}.
% % See \cref{app:rl} for more details of our experiments.
% We also compare to deep deterministic policy gradient (DDPG) \cite{lillicrapContinuousControlDeep2016}, a model-free RL algorithim (yellow).
% % As expected DDPG is the least sample inefficient
% The training curves in \cref{fig:rl} show that \our's uncertainty estimates are good for exploration as it converges in fewer episodes.

% We compare our method with deep deterministic policy gradient (DDPG) \cite{lillicrapContinuousControlDeep2016} -- a model-free RL baseline --
% which is sample inefficient.

 \todo{I'm thinking we should take out the SFR with updates result}
\begin{figure}[!t]
 \centering
 \begin{subfigure}[c]{.24\textwidth}
 \resizebox{\textwidth}{!}{%
 \begin{tikzpicture}[inner sep=0,outer sep=0]

   % Draw decorated 'ground'
   \draw[postaction={draw, decorate, decoration={border, angle=-45,
					amplitude=1cm, segment length=.5cm}}] (-3,-1.5) -- (12,-1.5);

   % The cart
   \draw[draw=black,fill=black!50,draw=black,line width=3pt,rounded corners=1mm] (0,0) rectangle (9cm,3cm);
   \node[fill=black,circle,minimum size=.5cm] (dot) at (4.5cm,3cm) {};

   % Wheels
   \node[fill=white,draw=black,line width=3pt,circle,minimum size=2cm,,fill=black!50] at (2cm,-.5cm) {};
   \node[fill=white,draw=black,line width=3pt,circle,minimum size=2cm,fill=black!50] at (7cm,-.5cm) {};

   % The arm
   \node[anchor=north,minimum width=1cm,minimum height=14cm,draw=black,rotate=-20,rounded corners=5mm,yshift=7mm,xshift=-.3mm,fill=white,draw=black,line width=3pt] at (dot) {};
   \node[fill=black,circle,minimum size=.5cm] at (dot) {};

   % Markings
   \draw[loosely dashed,line width=1pt] (4.5,6) -- (4.5,-10);

   % Arrow
   \draw[->,black,line width=3pt,-{Latex[length=7mm,width=7mm]}] (1,5) --node[above,outer sep=8pt]{\scalebox{4}{$x$}} (4.5,5);

   \def\centerarc[#1](#2)(#3:#4:#5)% Syntax: [draw options] (center) (initial angle:final angle:radius)
   { \draw[#1] ($(#2)+({#5*cos(#3)},{#5*sin(#3)})$) arc (#3:#4:#5); }

   % Draw arc
   \centerarc[black,line width=3pt](dot)(250:270:11)
   \node at (2.5,-9) {\scalebox{4}{$\theta$}};

   %\node[white,minimum size=1cm] at (4.5,-13) {};

 \end{tikzpicture}}
 \end{subfigure}
 \hfill
 \begin{subfigure}[c]{.74\textwidth}
   \centering\scriptsize
   \setlength{\figurewidth}{\textwidth}
   \setlength{\figureheight}{.5\figurewidth}
   \pgfplotsset{axis on top,ymajorgrids,axis line style={draw=none},legend style={at={(1,1)},anchor=north west}}
   \pgfplotsset{grid style={line width=.1pt, draw=gray!10,dashed},legend style={fill=white}}
   \input{./fig/rl.tex}
 \end{subfigure}
 \hfill
 \caption{\textbf{Cartpole swingup with sparse reward} Training curves showing that \our's uncertainty estimates improve sample efficiency in RL.
   Our method (blue/green) converges in fewer environment steps than the baseline model-based RL method (magenta) and DDPG, the model-free baseline (yellow).}
 \label{fig:rl}
\end{figure}

%

% \begin{figure}
%   \centering
%   \includegraphics[width=0.5\textwidth, angle=270]{fig/weight-space-to-functio-space.pdf}
%   \caption{}
% \end{figure}

% \begin{figure}
%   \centering
%   \includegraphics[width=0.5\textwidth, trim=0 100 0 10]{fig/cartpole-training-curves.pdf}
%   \caption{}
% \end{figure}

% \begin{table}
%   \caption{Negative test log likelihood (lower is better) on UCI classification tasks (2 hidden layers, 50 tanh). Our SVGP predictive outperforms the GLM predictive. }
% \end{table}

\section{Discussion and conclusion}
\label{sec:conclusion}
%
In this paper, we have introduced \our, a novel approach for representing neural networks in sparse function-space, exploiting the dual parameters for an efficient low-rank approximation that accommodates information from the entire data distribution. Our method offers a powerful mechanism for capturing predictive uncertainty, updating the model with new data without retraining, and providing a compact representation suitable for continual learning. These aspects were demonstrated in a wide range of problems, datasets, and learning contexts. We showcased our method's ability to capture uncertainty in UCI and image tasks under supervised learning (\cref{sec:uci}), demonstrated robustness on image data sets (\cref{,sec:image}), established its potential for continual learning (\cref{sec:cl-exp}), and finally, verified its applicability in reinforcement learning scenarios (\cref{sec:rl-exp}).

% Discussion points
It is worth noting that the NTK linearises the network around the MAP solution $\weights^{*}$,
so the function space prior (and thus posterior) is only a locally linear approximation.
In contrast to conventional GP kernels, the NTK does not have any hyperparameters which need to be learned.
It is also worth noting that the NTK may be highly non stationary and is dependent on the NN architecture, for example, the activation functions.

The broader impact of this work lies in its potential to provide tooling for how neural networks are utilized, offering more efficient and principled ways of handling uncertainty and continual learning. This contribution, we believe, has significant implications for future applications of machine learning in dynamic, real-world settings where data is unevenly distributed, uncertain, and continuously evolves.

A reference implementation of the methods presented in this paper is currently available as supplementary material and will be made available under the MIT License on GitHub upon acceptance.

%\section*{Broader Impact}

% \section*{References}
%\small
%\printbibliography
%\normalsize
% TODO make bibliography small a better way

%References follow the acknowledgments. Use unnumbered first-level heading for
%the references. Any choice of citation style is acceptable as long as you are
%consistent. It is permissible to reduce the font size to \verb+small+ (9 point)
%when listing the references.
%Note that the Reference section does not count towards the page limit.
%\medskip

\clearpage

\phantomsection%
\addcontentsline{toc}{section}{References}
\begingroup
\small
\bibliographystyle{abbrvnat}
\bibliography{bibliography}%zotero-library
\endgroup

\clearpage

\nipstitle{
    {\Large Supplementary Material:} \\
    Sparse Function-space Representation \\ of Neural Networks}
\pagestyle{empty}

\appendix

This supplementary document is organized as follows. 
%
\cref{app:method} provides a more extensive overview of the techncical details of \our.
%
\cref{app:cl} covers the sparse-functional regularisation (SFR) setup for continual learning.
%
\cref{app:rl} provides a more extensive writeup of the reinforcement learning setup used in the experiments.
%
\cref{app:experiments} provides the full details of each individual experiments in the main paper.

\section{Method details}
\label{app:method}
%
Optionally include extra information (complete proofs, additional experiments and plots) in the appendix.
This section will often be part of the supplemental material.

\section{\our for continual learning}
\label{app:cl}
\subsection{Derivation of the regularizer term}
An alternative view of the dual variables is that they parameterize approximate likelihoods, as shown in \citep{adam2021dual, khan2017conjugate}. We can rewrite the GP posterior for $q(\vu)$ in terms of the following approximate normal distributions:
\begin{equation}
 q(\vu) \propto \Norm(0, \MKzz) \prod_{i=1}^n \exp \! \left(-\frac{\tilde{\beta}_i}{2}(\tilde{y}_i - \va_i^\top \vu)^2 \right) = \Norm(0, \MKzz) \, \exp\left((\bar{\vy} - \vu)^\top \bar{\MSigma}^{-1}(\bar{\vy} - \vu)\right)  
\end{equation}
The first normals are the approximate likelihoods where $\va_i^\top = \vkzi^\top \MKzz^{-1}$ and $\tilde{y}_i = \hat{\alpha}_i + f_{\vw^*}(\vx_i)\hat{\beta}$,  replacing $\va_i^\top \vu$ with $\vx_i^{\top}\vw$ and we can see the similarity between the sparse GP and a linear model where the kernel matrix determines the prior. To arrive at the MVN Gaussian, we need to rearrange the quadratic terms to match the prior sufficient statistics we present a derivation in [appendix]. Performing conjugate summation of prior and likelihood, we would arrive at the sparse GP posterior at the inducing points. The correlated covariance structure $\MSigma$ and the corresponding $\tilde{\vy}$ are simply a different form of the sparse dual variables,
\begin{equation}
\quad \bar{\MSigma} =  \MKzz \vbeta_\vu^{-1} \MKzz \quad \bar{\vy} = \MKzz \vbeta_{\vu}\valpha_{\vu}.
\end{equation}
Written as above, we can see that the spare dual variables can be interpreted as a sparse MVN representation. We now show how we use this for a regularizer in continual learning with more details in [appendix].

\subsection{Extending the CL regularizer to multi-class settings}
\begin{equation}
	\bar{\MB}^{-1}_s = \MKzz^{-1} \vbeta_\vu \MKzz^{-1} \in \R^{C \times m \times m} \quad \MKzz \in \R^{C \times m \times m} 
\end{equation}

\begin{equation}
	\mathcal{R_\textit{SFR}}(\mathbf{w}) = \sum_{s=1}^{t-1}	\sum_{k \in 	C}\left[\left(f_{\vw, k}(\MZ_{s}) - f_{\vw_{s}, k}(\MZ_s) \right)^\T \bar{\MB}^{-1}_{s} \left(f_{\vw, k}(\MZ_{s}) - f_{\vw_{s, k}}(\MZ_s) \right) \right] 
\end{equation}

\section{Sparse functional model-based reinforcement learning}
\label{app:rl}

We consider environments with states \(\state \in \stateDomain \subseteq \R^{D_{\state}} \),
actions \(\action \in \actionDomain \subseteq \R^{D_{\action}}\) and transition dynamics
\(\transitionFn: \stateDomain \times \actionDomain \rightarrow \stateDomain \), such that
$\state_{t+1} = \transitionFn(\state_{t}, \action_{t}) + \noise_{t}$, where  $\noise_{t}$
is i.i.d. transition noise.
We consider the episodic setting where the system is reset to an initial state $\state_{0}$ at each episode and we
assume that there is a known reward function $r : \stateDomain \times \actionDomain \rightarrow \R$.
% Following the Markov decision process formulation \cite{bellmanMarkovianDecisionProcess1957a}, we
% denote the states \(\state \in \stateDomain \subseteq \R^{D_{\state}} \) and actions \(\action \in \actionDomain \subseteq \R^{D_{\action}}\)
% of the sytem, the reward function $r : \stateDomain \times \actionDomain \rightarrow \R$, and transition dynamics
% \(\transitionFn: \stateDomain \times \actionDomain \rightarrow \stateDomain \), such that
% $\state_{t+1} = \transitionFn(\state_{t}, \action_{t}) + \noise_{t}$ where  $\noise_{t}$
% is i.i.d. transition noise.
% $\mathbb{E}_{\noise_{0:\infty}} \big[ \sum_{t=0}^{\infty} \discount^{t} \rewardFn(\state_{t},\action_{t}) \big]$
% \begin{align} \label{eq-model-free-objective}
% \policy^{*} = \arg \max_{\policy \in \policyDomain} J(\transitionFn, \policy) = \arg \max_{\policy \in \policyDomain} \mathbb{E}_{\noise_{0:\infty}} \bigg[ \sum_{t=0}^{\infty} \discount^{t} \rewardFn(\state_{t},\action_{t}) \bigg]
% \quad \text{s.t. } \state_{t+1} = \transitionFn(\state_{t}, \action_{t}) + \noise_{t},
% \end{align}
The goal of RL is to find the policy \(\pi : \stateDomain \rightarrow \actionDomain\)
(from a set of policies $\Pi$) that maximises the sum of discounted rewards
in expectation over the transition noise,
\begin{align} \label{eq-model-free-objective}
J(\transitionFn, \policy) = \mathbb{E}_{\noise_{0:\infty}} \bigg[ \sum_{t=0}^{\infty} \discount^{t} \rewardFn(\state_{t},\action_{t}) \bigg]
\quad \text{s.t. } \state_{t+1} = \transitionFn(\state_{t}, \action_{t}) + \noise_{t},
\end{align}
where $\gamma \in [0, 1)$ is a discount factor.
In this work, we consider model-based RL where a model of the transition dynamics is learned \(f_{\mathbf{w}} \approx \transitionFn\) and then used by a planning algorithm.
A simple approach is to use the learned dynamic model $f_{\mathbf{w}^{*}}$ and maximise the objective in \cref{eq-model-free-objective},
\begin{align} \label{eq-greedy}
  \policy^{\text{Greedy}} &= \arg \max_{\pi \in \Pi} J(f_{\mathbf{w}^{*}}, \pi).
\end{align}
However, we can leverage the method in \cref{sec:methods} to obtain a function-space posterior over the learned dynamics $q_{\mathbf{u}}(\hat{\transitionFn} \mid \dataset)$,
where $\mathcal{D}$ represents the state transition data set \(\mathcal{D} = \{(s_{i},a_{i}), s_{i+1}\}_{i=0}^{N}\).
Importantly, the uncertainty represented by this posterior distribution can be used to balance the exploration-exploitation trade-off,
using approaches such as posterior sampling \cite{osbandWhyPosteriorSampling2017,osbandMoreEfficientReinforcement2013},
\begin{align} \label{eq-posterior-sampling}
  \policy^{\text{PS}} &= \arg \max_{\pi \in \Pi} J(\hat{f}, \pi)
\quad \text{s.t. } \tilde{\transitionFn} \sim q_{\mathbf{u}}(\hat{\transitionFn} \mid \dataset),
\end{align}
where a function $\tilde{\transitionFn}$ is sampled from the (approximate) posterior $q_{\mathbf{u}}(\hat{\transitionFn} \mid \dataset)$ and used to find a policy as
in \cref{eq-greedy}.
Intuitively, this strategy will explore where the model has high uncertainty, which in turn will reduce the model's uncertainty as data is collected and used to
train the model.

\todo{could try to implement Pathwise conditioning in function-space}

Model Predictive Path Integral (MPPI) control
\cite{panSample2015}
\cite{williamsModel2017}
\todo{what is correct citation for MPPI?}

% \cref{alg-mbrl} shows the typical model-based RL loop.
\begin{align} \label{eq-fast-update-mpc}
  \policy_{i+1}^{\text{PS}}(\state) = \arg \max_{\action_{0}} \max_{\action_{1:\Horizon}}
\E \bigg[ \sum_{t=0}^{H-1} \gamma^{t} r(\state_{t},\action_{t}) \mid \state_{0}=\state  \bigg] + Q_{\theta}(\state_{\Horizon}, \action_{H})
\quad \text{s.t. } \state_{t+1} &= \hat{\transitionFn}(\state_{t}, \action_{t}) + \noise_{t}
\end{align}

with $\hat{\transitionFn} \sim p(\transitionFn \mid \dataset)$

We use deep deterministic policy gradient (DDPG) \cite{lillicrapContinuousControlDeep2016} to learn an action value function $Q_{\theta}$.
Note that we also learn a policy but its sole purpose is for learning the value function.

\subsection{Experiment Configuration}
This section details how we configured and ran our reinforcement learning experiments.

\textbf{Dynamic model}
In all experiments we used an MLP dynamic model with a single hidden layer of width 64 and TanH activation functions.
At each episode we used Adam \cite{adam} to optimise the NN parameters for $5000$ iterations with a learning rate of $0.001$.
We reset the optimizer after each episode.
As we are performing regression we instantiate the loss function in \cref{eq-empirical-risk} as the well-known mean squared error.
This corresponds to a Gaussian likelihood with unit variance.
We then set the prior precision as $\delta=0.0001$.
% It is worth noting that $\delta$ effects both the neur
In all experiments our sparse function-space representation uses $m=128$ inducing points and this seemed to be sufficient.

\textbf{Model predictive path integral (MPPI)}
MPPI is an online planning algorithim which iteratively improves the action trajectory $\action_{t:t+H}$ using samples.
At each iteration $j$, $N=256$ trajectories are sampled according to the currect action trajectory $\action^{j}_{t:t+H}$.
The $K=32$ top trajectories with highest returns $\sum_{h=0}^{H} = r(\state^{j}_{t+h}, \action^{j}_{t+h})$ are selected.
The next action trajectory $\action^{j+1}_{t:t+H}$ is then computed by taking the weighted average of the top $K=32$ trajectories
with weights from the softmax over returns from top $K=32$ trajectories.

$\gamma 0.9$
$\tau=0.005$
horizon $H=5$
temperature $0.5$
momentum $0.1$

\textbf{Initial data set}
We collect an initial data set using a random policy for one episode.

\textbf{DDPG}
action value function is an MLP with a single hidden layer of width $128$ with ELU activation functions.
We train the DDPG agent using Adam for $500$ iterations at each episode, using a learning rate $0.0001$.

\section{Experiment details}
\label{app:experiments}

\subsection{Uncertainty quantification on UCI data sets}
\label{app:uci}

\subsection{Uncertainty benchmarks for image data}
\label{app:image}

\subsection{\our for continual learning}
\label{app:cl-experiment}
\section{Extending the CL regularizer to multi-class settings}
\label{sec:cl_multioutput}
\begin{equation}
	\bar{\MB}^{-1}_s = \MKzz^{-1} \vbeta_\vu \MKzz^{-1} \in \R^{C \times m \times m} \quad \MKzz \in \R^{C \times m \times m} 
\end{equation}

\begin{equation}
	\mathcal{R_\textit{SFR}}(\mathbf{w}) = \sum_{s=1}^{t-1}	\sum_{k \in 	C}\left[\left(f_{\vw, k}(\MZ_{s}) - f_{\vw_{s}, k}(\MZ_s) \right)^\T \bar{\MB}^{-1}_{s} \left(f_{\vw, k}(\MZ_{s}) - f_{\vw_{s, k}}(\MZ_s) \right) \right] 
\end{equation}

\subsection{Reinforcement learning experiment}
\label{app:rl-experiment}

%\begin{table}[h]
%    \centering
%    \input{./tables/example_table.tex}
%\end{table}

\subsection{Biblatex}
Rember when using biblatex to use 'parencite' for \citep{kamtheDataEfficient2018} and when using natbib to use 'citep'.

%\bibliography{biblio.bib}
\end{document}